---
title: 'CKME 136 X10-DATA ANALYTICS: CAPSTONE PROJECT'
author: "Ekaete Czub"
date: February 10, 2019
output:
  word_document: default
  html_document: default
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
## knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(echo = TRUE)
```

##Installation of relevant packages.

```{r}
#install.packages("ISwR")
#install.packages("caret")
#install.packages("MASS")
#install.packages("lubridate")
#install.packages("stringr")
#install.packages("corrplot")
#install.packages("ggcorrplot")
#install.packages("Metrics")
library(ISwR)
library(caret)
library(MASS)
library(lubridate)
library(stringr)
library(corrplot)   #### <---------------  Added for correlation plot
library(ggcorrplot) #### <---------------  Added for correlation plot
library(Metrics)
```

## Read "Avocado.csv file" from folder or website:

```{r}
Avocado <- read.csv(file.choose(), header = T, sep = ",", stringsAsFactors = F, na.strings = c(NA,"","?"))
avobk <- Avocado
```

## Cleaning data:correcting column names
```{r}
#names(Avocado)  
#names(Avocado)[1]<- "No"
names(Avocado)[names(Avocado) =="X"]<- "No"
names(Avocado)[names(Avocado) =="X4046"]<-"PLU4046"
names(Avocado)[names(Avocado) =="X4225"]<-"PLU4225"
names(Avocado)[names(Avocado) =="X4770"]<-"PLU4770"
#head(Avocado)
str(Avocado)
any(is.na(Avocado))
```

#column: "No"  - Categorical column  <--- This column will be removed .....
```{r}
## Removing column 1(No)
Avocado<- Avocado[, -1] 
str(Avocado)
```

# column:Date - Categorical column
```{r}
### names(Avocado)[2]
names(Avocado)[1]  ### <----- Date becomes the first column because the first column is removed..
### Converting the date column from character to date
Avo_Date <- Avocado$Date
Avo_Date <- as.Date(Avo_Date, format= "%Y-%m-%d")
Avocado$Date <- Avo_Date
Avocado$Month <- month(Avocado$Date)
Avocado$Month <- as.factor(Avocado$Month)
Avocado$Day   <- day(Avocado$Date)
Avocado$Day   <- as.factor(Avocado$Day)
Avocado$year  <- as.factor(Avocado$year)
##Avocado$Week  <- as.factor(week(Avocado$Date))
Avocado <- Avocado[-1]
```

## column: AveragePrice - Numeric column
```{r}
names(Avocado)[1]
Avo_AvgPrice <- Avocado$AveragePrice
summary(Avo_AvgPrice)
hist(Avo_AvgPrice, breaks = 20, main="Histogram: Avocado Price distribution")###  ----> This goes to the INITIAL analysis <---
boxplot(Avo_AvgPrice, main = "Boxplot of Avocado Average Price with outliers", ylab="count",horizontal = F, outcol = "red", col="red")  ###  ----> This goes to the INITIAL analysis <---
pout <- boxplot.stats(Avo_AvgPrice)$out
while (length(pout)>0) {
  outrecs <- which(Avo_AvgPrice %in% pout)
  Avo_AvgPrice[outrecs] <- NA
  Avo_AvgPrice[is.na(Avo_AvgPrice)] <- median(Avo_AvgPrice, na.rm = T)
  pout <- boxplot.stats(Avo_AvgPrice)$out
}
Avocado$AveragePrice<-Avo_AvgPrice
hist(Avocado$AveragePrice, main= "Histogram: Avocado Average Price distribution",col = "red") ###  ----> This goes to the FINAL analysis <---
boxplot(Avocado$AveragePrice, main ="Boxplot of Avocado Average Price", ylab="count",xlab="Average Price", col="red") ### ----> This goes to the FINAL analysis <---
summary(Avo_AvgPrice)
```

# column: Total.Volume - Numeric column  
```{r}
names(Avocado)[2]
Avo_Total.Vol <- Avocado$Total.Volume
summary(Avo_Total.Vol) 
hist(Avo_Total.Vol, main= "Histogram: Total Volume distribution of Avocado", cex.main=1, ylab = "number of Avocado sold",cex.lab=1)###  ----> This goes to the INITIAL analysis <---
boxplot(Avo_Total.Vol, main = "Boxplot of Total Volume of Avocado with outliers", cex.main=1, ylab = "count",cex.axis=1, xlab="Total Volume", cex.lab=1, horizontal = F, outcol = "orange", col= "orange") ###  ----> This goes to the INITIAL analysis <---
poutTV<-boxplot.stats(Avo_Total.Vol)$out
while (length(poutTV)>0) {
  outrecsTV <- which(Avo_Total.Vol %in% poutTV)
  Avo_Total.Vol[outrecsTV] <- NA
  Avo_Total.Vol[is.na(Avo_Total.Vol)] <- median(Avo_Total.Vol, na.rm = T)
  poutTV <- boxplot.stats(Avo_Total.Vol)$out
}
Avocado$Total.Volume <- Avo_Total.Vol
hist(Avocado$Total.Volume, main= "Histogram: Total Volume distribution of Avocado",cex.main=1,ylab= "number of Avocado sold",cex.axis=1, col= "orange") ### ----> This goes to the FINAL analysis <---
boxplot(Avocado$Total.Volume, main = "Boxplot of Total Volume of Avocado", cex.main=1, ylab="count",cex.axis=1, xlab = "Total Volume",cex.lab=1, col="orange") ### ----> This goes to the FINAL analysis <---
summary(Avo_Total.Vol)
```

# column: PLU4046- numeric column 
```{r}
names(Avocado)[3]
Avo_PLU4046 <- Avocado$PLU4046
summary(Avo_PLU4046) 
hist(Avo_PLU4046, main = "Histogram of Avocado Product Look code(PLU) 4046", cex.main=1, ylab = "Number of Avocado with PLU4046 sold", cex.axis=1)###  ----> This goes to the INITIAL analysis <---
boxplot(Avo_PLU4046, main = "Boxplot of Avocado PLU4046 with outliers", cex.main=1,  ylab="count", cex.axis=1,xlab= "Avocado PLU4046", cex.lab=1,horizontal = F, outcol="yellow", col="yellow")###  ----> This goes to the INITIAL analysis <---
poutC1<-boxplot.stats(Avo_PLU4046)$out
while (length(poutC1)>0) {
  outrecsC1 <- which(Avo_PLU4046 %in% poutC1)
  Avo_PLU4046[outrecsC1] <- NA
  Avo_PLU4046[is.na(Avo_PLU4046)] <- median(Avo_PLU4046, na.rm = T)
  poutC1 <- boxplot.stats(Avo_PLU4046)$out
}
Avocado$PLU4046<-Avo_PLU4046
hist(Avocado$PLU4046,main= "Histogram of Avocado PLU4046", cex.main=1,ylab= "number of Avocado with PLU4046 sold",cex.axis=1, col="yellow")### ----> This goes to the FINAL analysis <---
boxplot(Avocado$PLU4046, main="Boxplot of Avocado PLU4046", cex.main=1,ylab="count",cex.axis=1, xlab="Avocado PLU4046", cex.lab=1,col = "yellow")### ----> This goes to the FINAL analysis <---
summary(Avo_PLU4046)
```

# column: PLU4225 - numeric column
```{r}
names(Avocado)[4]
Avo_PLU4225 <- Avocado$PLU4225
summary(Avo_PLU4225) 
hist(Avo_PLU4225, main="Distribution of Avocado PLU4225", ylab="number of Avocado with PLU4225 sold")###  ----> This goes to the INITIAL analysis <---
boxplot(Avo_PLU4225, main = "Boxplot of Avocado PLU4225 with outliers",ylab="count", xlab= "Avocado PLU4225", horizontal = F, outcol="pink", col="pink")###  ----> This goes to the INITIAL analysis <---
poutC2<-boxplot.stats(Avo_PLU4225)$out
while (length(poutC2)>0) {
  outrecsC2 <- which(Avo_PLU4225 %in% poutC2)
  Avo_PLU4225[outrecsC2] <- NA
  Avo_PLU4225[is.na(Avo_PLU4225)] <- median(Avo_PLU4225, na.rm = T)
  poutC2 <- boxplot.stats(Avo_PLU4225)$out
}
Avocado$PLU4225<-Avo_PLU4225
hist(Avocado$PLU4225, main="Distribution of Avocado PLU4225", ylab= "number of Avocado with PLU4225 sold",col="pink")### ----> This goes to the FINAL analysis <---
boxplot(Avocado$PLU4225, main="Boxplot of Avocado PLU4225", ylab="count", xlab= "Avocado PLU4225", col="pink")### ----> This goes to the FINAL analysis <---
summary(Avo_PLU4225)
```

# column: PLU4770 - numeric column
```{r}
names(Avocado)[5]
Avo_PLU4770 <- Avocado$PLU4770
summary(Avo_PLU4770) 
hist(Avo_PLU4770, main="Distribution of Avocado PLU4770", ylab="number of Avocado with PLU4770 sold")###  ----> This goes to the INITIAL analysis <---
boxplot(Avo_PLU4770, main = "Boxplot of Avocado PLU4770 with outliers", xlab= "Avocado PLU4770", horizontal = F, outcol="green", col="green")###  ----> This goes to the INITIAL analysis <---
poutC3<-boxplot.stats(Avo_PLU4770)$out
while (length(poutC3)>0) {
  outrecsC3 <- which(Avo_PLU4770 %in% poutC3)
  Avo_PLU4770[outrecsC3] <- NA
  Avo_PLU4770[is.na(Avo_PLU4770)] <- median(Avo_PLU4770, na.rm = T)
  poutC3 <- boxplot.stats(Avo_PLU4770)$out
}
Avocado$PLU4770<-Avo_PLU4770
hist(Avocado$PLU4770, main="Distribution of Avocado PLU4770", ylab="number if Avocado with PLU4770 sold",col="green")### ----> This goes to the FINAL analysis <---
boxplot(Avocado$PLU4770, main = "Boxplot of Avocado PLU 4770", ylab="count",xlab="Avocado PLU4770", col="green")### ----> This goes to the FINAL analysis <---
summary(Avo_PLU4770)
```

# column: Total.Bags- numeric column
```{r}
names(Avocado)[6]
Avo_Total.Bags <- Avocado$Total.Bags
summary(Avo_Total.Bags) 
hist(Avo_Total.Bags, main="Histogram of Total Bags of Avocado", ylab="number of Total Bags of Avocado")###  ----> This goes to the INITIAL analysis <---
boxplot(Avo_Total.Bags, main = "Boxplot of Total Bags of Avocado with outliers", ylab="count", xlab="Total Bags", horizontal = F, outcol = "purple", col="purple")###  ----> This goes to the INITIAL analysis <---
poutTB<-boxplot.stats(Avo_Total.Bags)$out
while (length(poutTB)>0) {
  outrecsTB <- which(Avo_Total.Bags %in% poutTB)
  Avo_Total.Bags[outrecsTB] <- NA
  Avo_Total.Bags[is.na(Avo_Total.Bags)] <- median(Avo_Total.Bags, na.rm = T)
  poutTB <- boxplot.stats(Avo_Total.Bags)$out
}
Avocado$Total.Bags<-Avo_Total.Bags
hist(Avocado$Total.Bags, main="Histogram of Total Bags of Avocado", ylab="number of Total Bags of Avocado",col = "purple")### ----> This goes to the FINAL analysis <---
boxplot(Avocado$Total.Bags, main="Boxplot of Total Bags of Avocado", ylab="count", xlab="Total Bags", col="purple")### ----> This goes to the FINAL analysis <---
summary(Avo_Total.Bags)
```

# column: Small.Bags- numeric column
```{r}

names(Avocado)[7]
Avo_Small.Bags <- Avocado$Small.Bags
summary(Avo_Small.Bags) 
hist(Avo_Small.Bags, main="Histogram of Small Bags of Avocado", ylab="number of Small Bags of Avocado")###  ----> This goes to the INITIAL analysis <---
boxplot(Avo_Small.Bags, main = "Boxplot of Small Bags of Avocado with outliers",ylab="count", xlab="Small Bags", horizontal = F, outcol = "brown", col="brown")###  ----> This goes to the INITIAL analysis <---
poutSB<-boxplot.stats(Avo_Small.Bags)$out
while (length(poutSB)>0) {
  outrecsSB <- which(Avo_Small.Bags %in% poutSB)
  Avo_Small.Bags[outrecsSB] <- NA
  Avo_Small.Bags[is.na(Avo_Small.Bags)] <- median(Avo_Small.Bags, na.rm = T)
  poutSB <- boxplot.stats(Avo_Small.Bags)$out
}
Avocado$Small.Bags<-Avo_Small.Bags
hist(Avocado$Small.Bags, main="Histogram of Small Bags of Avocado" , ylab="number of Small Bags of Avocado", col="brown")### ----> This goes to the FINAL analysis <---
boxplot(Avocado$Small.Bags, main="Boxplot of Small Bags of Avocado", ylab="count", xlab="Small Bags", col="brown")### ----> This goes to the FINAL analysis <---
summary(Avo_Small.Bags)
```

# column: Large.Bags - numeric column
```{r}
names(Avocado)[8]
Avo_Large.Bags <- Avocado$Large.Bags
summary(Avo_Large.Bags) 
hist(Avo_Large.Bags, main="Histogram of Large Bags of Avocado", ylab="number of Large Bags of Avocado")###  ----> This goes to the INITIAL analysis <---
boxplot(Avo_Large.Bags, main = "Boxplot of Large Bags of Avocado with outliers", ylab="count", xlab="Large Bags",horizontal = F, outcol = "gold", col="gold")###  ----> This goes to the INITIAL analysis <---
poutLB<-boxplot.stats(Avo_Large.Bags)$out
while (length(poutLB)>0) {
  outrecsLB <- which(Avo_Large.Bags %in% poutLB)
  Avo_Large.Bags[outrecsLB] <- NA
  Avo_Large.Bags[is.na(Avo_Large.Bags)] <- median(Avo_Large.Bags, na.rm = T)
  poutLB <- boxplot.stats(Avo_Large.Bags)$out
}
Avocado$Large.Bags<-Avo_Large.Bags
hist(Avocado$Large.Bags, main="Histogram of Large Bags of Avocado", ylab="number of Large Bags of Avocado", col = "gold")### ----> This goes to the FINAL analysis <---
boxplot(Avocado$Large.Bags, main="Histogram of Large Bags of Avocado", ylab="count", xlab="Large Bags", col="gold")### ----> This goes to the FINAL analysis <---
summary(Avo_Large.Bags)
```

# column: XLarge.Bags - numeric column  ---->  THIS COLUMN WILL BE REMOVED <-----

```{r}
names(Avocado)[9]
Avo_XLarge.Bags <- Avocado$XLarge.Bags
summary(Avo_XLarge.Bags) 
hist(Avo_XLarge.Bags)###  ----> This goes to the INITIAL analysis <---
boxplot(Avo_XLarge.Bags, main = "Avo_XLarge.Bags with outliers", horizontal = F, outcol = "blue", col="blue")###  ----> This goes to the INITIAL analysis <---
poutXLB<-boxplot.stats(Avo_XLarge.Bags)$out
while (length(poutXLB)>0) {
  outrecsXLB <- which(Avo_XLarge.Bags %in% poutXLB)
  Avo_XLarge.Bags[outrecsXLB] <- NA
  Avo_XLarge.Bags[is.na(Avo_XLarge.Bags)] <- median(Avo_XLarge.Bags, na.rm = T)
  poutXLB <- boxplot.stats(Avo_XLarge.Bags)$out
}
Avocado$XLarge.Bags<-Avo_XLarge.Bags
hist(Avocado$XLarge.Bags)
boxplot(Avocado$XLarge.Bags, main="Extra Large Bags of Avocado")
#Removing column 11(XLarge.Bags)
Avocado$Total.Bags <- Avocado$Small.Bags + Avocado$Large.Bags
Avocado<- Avocado[, -9]
## str(Avocado)
```
### ---------------------------------------------------------------------------------
### --->  COLUMN NUMBERS CHANGED AGAIN AFTER REMOVING the column XLarge.Bags <-------
### ---------------------------------------------------------------------------------
# column: type - categorical column
```{r}
names(Avocado)[9]
Avo_type <- Avocado$type
table(Avo_type)
type<-table(Avo_type)
barplot(type, main = "Barplot of Avocado Types", col=rainbow(2))
```

# column: year - categorical column
```{r}
# names(Avocado)[10]
# Avo_year <- Avocado$year
# Avo_year <- factor(Avo_year)
#head(Avo_year)
Years <- table(Avocado$year)
barplot(Years, main = "Avocado Years Covarage", col=rainbow(4))
```

# column: region - categorical column
```{r}

names(Avocado)[11]
Avo_region<- Avocado$region
region<-table(Avo_region, useNA = 'ifany')
region

barplot(region[1:10],  main = "Barplot of Avocado Regions")
barplot(region[11:20], main = "Barplot of Avocado Regions")
barplot(region[21:30], main = "Barplot of Avocado Regions")
barplot(region[31:40], main = "Barplot of Avocado Regions")
barplot(region[41:50], main = "Barplot of Avocado Regions")
barplot(region[51:length(region)], main = "Barplot of Avocado Regions")
```

#correlation between numeric columns
```{r}
corr_df<-Avocado[1:8]  
## plot(d)  <--- Imaage is saved. 
corr_mtrx <- cor(corr_df)
corrplot(corr_mtrx)
ggcorrplot(corr_mtrx)
ggcorrplot(corr_mtrx, lab = T, hc.order = T, type = 'upper')
```

#Splitting dataset to 70% training and 30% test sets
```{r}
Avocado$type <- as.factor(Avocado$type)
Avocado$region <- as.factor(Avocado$region)
Avocado_Train <-sample(nrow(Avocado), floor(nrow(Avocado)*0.7))
train_data <-Avocado[Avocado_Train,]
test_data  <-Avocado[-Avocado_Train,]
# Building the Multiple Linear Regression Model
set.seed(123)
Avocado_Model <-lm(AveragePrice~., data=train_data)
summary(Avocado_Model)
prediction_prices<-predict(Avocado_Model, interval="prediction", newdata=test_data)
summary(prediction_prices)
```
```{r}
red_Avocado <- Avocado[,c("AveragePrice","Total.Volume","PLU4046","PLU4225","PLU4770","type")]
plot(red_Avocado)
avo_lm_rd <- lm(AveragePrice~Total.Volume+PLU4046+PLU4225+PLU4770+type, data = train_data)
summary(avo_lm_rd)
pric_pred_rd <- predict(Avocado_Model, interval="prediction", newdata=test_data)
summary(avo_lm_rd)
anova(avo_lm_rd, Avocado_Model)
```

#Plotting errors on a histogram
```{r}
#errors<-prediction_prices[,"fit"]-test_data$AveragePrice
#errors
#hist(errors)

